### 2018-11-03 experiments.tex
getting ready for TR

We introduced a subtle bug into {\tt pyfakefs}, where the {\tt remove}
call checks that its target is not a directory, and returns the
correct error, but still carries out the remove operation.  Using {\tt
  os.remove} to delete directories does not break any file system
invariants, but violates the Python {\tt os} specification (and,
indirectly, the usual POSIX implementation behavior where {\tt unlink}
does not work for directories).  Detecting this bug using the TSTL
{\tt pyfakefs} harness is normally impossible without using another
file system as a reference.  However, the fault was detected
immediately, even without using a reference, when we compiled the harness
with the {\tt --checkFailureDeterminism} flag.  Because vertical
reduction does not require running complete tests multiple times, and
does not affect the delta-debugging algorithm's performance, reducing
the failing test to 3 steps required less than a second.
Moreover, the overhead for the check for failure determinism in a
version of the code without the {\tt remove} error was
less than 8\%.  Detecting the fault using a reference file system 
required 17\% more testing time before detection, and took over twice as
long to reduce the failure to a
slightly longer failing test, which did not have {\tt remove} as its
final operation (since further operations are required to 
expose the bad file system state the operation introduces).

@@

We wanted to determine whether there were also simple faults that
could be detected by failure nondeterminism, but \emph{not} detected
by differential testing, and quantify the additional specification
strength provided by failure determinism checks.  We therefore used
{\tt universalmutator} \cite{RegExpMut} to produce 2,350 mutants of
the {\tt pyfakefs} core file system code.   We restricted the generation
to only mutate code covered during a 60 second run of the test
harness, with differential testing turned off.  In this
\emph{non-differential} mode, the harness will only detect a fault when it
causes an unexpected exception to be raised, or results in a timeout
due to, e.g., and infinite loop.

We first analyzed the mutants using 60 seconds of non-differential
testing, then with the full differential test plus
an additional 60 seconds of testing (to account for the observed 17\%
extra time to detect in the manually constructed example).   For both analyses, we tested
the surviving mutants using the non-differential harness for 60
seconds, but with failure nondeterminism checking.
We used the same random seed for
all runs, so that the only differences would be specification-based.

Non-differential testing, without failure determinism checks, killed
872 mutants, a 37.1\% kill rate.  Adding a failure determinism check
allowed the testing to kill an additional 98 mutants, an 11\% improvement.
The differential harness, with an additional 60 seconds of testing
time, killed 1148 mutants, improving the kill rate to 48.8\%.  Failure determinism added an
additional 71 mutants to the total killed, a 6\% improvement even for this complete
harness with a larger test budget.  The kill rates are generally low
because {\tt universalmutator}, by design, includes many operations
that can produce equivalent mutants, but can also produce hard-to-kill
non-equivalent mutants not produced by other mutation tools, e.g. code
that throws away exceptions raised by a function call, or code
reversing a list.

In order to further investigate the additional specification power
provided by failure nondeterminism detection, we inspected the mutants
killed by the failure determinism check but not by differential
testing.  The largest category of mutants not killed (22 of the 71 mutants) was what we refer
to as ``exception swallowing'' mutants, which transform a Python
statement into the same statement, but wrapped in a {\tt try} block
with a catch that ignores any raised exceptions, e.g., {\tt foo()} becomes:

@@

\noindent It is easy to
see that such mutants may introduce faults in the handling of
errors, and thus would tend to cause failure nondeterminism.  However,
these are a minority of the mutations introducing failure
nondeterminism.  Other mutation operators resulting in subtle flaws
not (at least easily) detectable by differential testing compared to a
correct file system include:  arithmetic operation changes, statement
deletions, logical operator modifications, constant replacements
(including replacement of a string with the empty string), and
introducing a break into a loop.

We also checked the cost of introducing failure determinism checking
by analyzing mutants that could be killed by both methods.  Even
though in some cases the failure determinism check allows a mutant to
be detected sooner, the mean time to kill mutants was 0.006 seconds
larger with failure determinism checking, and this change was significant by Wilcoxon test
$(p < 1 \times 10^{-15})$.

>>>>

We introduced a subtle bug into {\tt pyfakefs}, where the {\tt remove}
call checks that its target is not a directory, and returns the
correct error, but still carries out the remove operation.  Using {\tt
  os.remove} to delete directories does not break any file system
invariants, but violates the Python {\tt os} specification (and,
indirectly, the usual POSIX implementation behavior where {\tt unlink}
does not work for directories).  Detecting this bug using the TSTL
{\tt pyfakefs} harness is normally impossible without using another
file system as a reference.  However, the fault was detected
immediately, even without using a reference, when we compiled the harness
with the {\tt --checkFailureDeterminism} flag.  Because vertical
reduction does not require running complete tests multiple times, and
does not affect the delta-debugging algorithm's performance, reducing
the failing test to 3 steps required less than a second.
Moreover, the overhead for the check for failure determinism in a
version of the code without the {\tt remove} error was
less than 8\%.  Detecting the fault using a reference file system 
required 17\% more testing time before detection, and took over twice as
long to reduce the failure to a
slightly longer failing test, which did not have {\tt remove} as its
final operation (since further operations are required to 
expose the bad file system state the operation introduces).  For this
hypothetical subtle bug, failure determinism checks not only make it
possible to detect the fault without a reference implementation, it
improves on detection speed and ease of debugging even compared to a
full-blown reference implementation (the MacOS file system, operating
on a RAM disk) and strong, hand-tuned, differential testing.

@@

We wanted to determine whether there were also simple faults that
could be detected by failure nondeterminism, but \emph{not} detected
by differential testing, and quantify the additional specification
strength provided by failure determinism checks.  We therefore used
{\tt universalmutator} \cite{RegExpMut} to produce 2,350 mutants of
the {\tt pyfakefs} core file system code.   We restricted the generation
to only mutate code covered during a 60 second run of the test
harness, with differential testing turned off.  In this
\emph{non-differential} mode, the harness will only detect a fault when it
causes an unexpected exception to be raised, or results in a timeout
due to, e.g., an infinite loop.

We first analyzed the mutants using 60 seconds of non-differential
testing, then with the full differential test plus
an additional 60 seconds of testing (to make sure we accounted for the observed 17\%
extra time to detect in the manually constructed example: we want to
maximize the chance to detect a bug for the methods we compare
against).   For both comparisons, we then tested
the surviving mutants using the non-differential harness for 60
seconds, but with failure nondeterminism checking turned on.
We used the same random seed for
all runs, so that the only differences would be specification-based.

Non-differential testing, without failure determinism checks, killed
872 mutants, a 37.1\% kill rate.  Adding a failure determinism check
allowed the testing to kill an additional 98 mutants, an 11\% improvement.
The differential harness, with an additional 60 seconds of testing
time, killed 1148 mutants, improving the kill rate to 48.8\%.  Failure determinism added an
additional 71 mutants to the total killed, a 6\% improvement even for
this strong differential harness with a larger test budget.  The kill rates are generally low
because {\tt universalmutator}, by design, includes many operations
that can produce equivalent mutants, but can also produce hard-to-kill
non-equivalent mutants not produced by other mutation tools, e.g. code
that throws away exceptions raised by a function call, or code
reversing a list.

In order to further investigate the additional specification power
provided by failure nondeterminism detection, we inspected the mutants
killed using the failure determinism check but not killed by the
strong differential testing.  The largest category of mutants not
killed (22 of the 71 mutants) was what we refer to as ``exception
swallowing'' mutants, which transform a Python statement into the same
statement, but wrapped in a {\tt try} block with a catch that ignores
any raised exceptions, e.g., {\tt foo()} becomes:

@@

\noindent It is easy to
see that such mutants may introduce faults in the handling of
errors, and thus would tend to cause failure nondeterminism.  However,
these are a minority of the mutations introducing hard-to-detect but
non-equivalent failure
nondeterminism.  Other mutation operators resulting in subtle flaws
not (at least easily) detectable by differential testing compared to a
correct file system include:  arithmetic operation changes, statement
deletions, logical operator modifications, constant replacements
(including replacement of a string with the empty string), and
introducing a break into a loop.  The variety of mutant types suggests
that no more specifically tailored strategy such as checking for
exception propagation, will work as well as introducing a notion of
failure nondeterminism.

We also checked the cost of introducing failure determinism checking
by analyzing mutants that could be killed by both methods.  Even
though in some cases the failure determinism check allows a mutant to
be detected sooner, the mean time to kill mutants was 0.006 seconds
larger with failure determinism checking, and this change was significant by Wilcoxon test
$(p < 1 \times 10^{-15})$.  While significant, this cost is probably
too small to be of any real practical importance.

### 2018-11-03 implementation.tex
getting ready for TR

@@

We therefore instead made nondeterminism detection a first-class
property in TSTL, using TSTL's existing notation for marking some
types of values as \emph{opaque} (not usefully compared for equality),
used in automatic abstraction-based testing.  In TSTL, a test is a
sequence of \emph{actions}; parameter values to method and function
calls of the Software Under Test (SUT) are stored in finite-sized variable \emph{pools}.  A
TSTL test is (essentially) a sequence of assignments to pool values,
and method/function calls using pool values (including method calls on
pool objects).  This formalism is a widely used and efficient way to
represent unit tests \cite{Pacheco,AndrewsTR}.  In TSTL, \emph{visible
value determinism} is based on comparing the values of \emph{all pool
variables} after each test action (pool values other than the one
assigned to must also be compared, since a common source of
nondeterminism is when a call results in a change to a value passed as
a parameter, or even due to a shared reference held by two values).
TSTL omits comparison only of pools marked as {\tt OPAQUE} in the
TSTL language test harness (and values that do not support equality
checking).  \emph{Final State Determinism} simply performs the same
comparison, but only on the final values of all pool variables (or a
set of designated pools) after a test has
finished executing.

>>>>

\subsection{Implementing Horizontal Nondeterminism Detection}

@@

We therefore instead made horizontal nondeterminism detection a first-class
property in TSTL, using TSTL's existing notation for marking some
types of values as \emph{opaque} (not usefully compared for equality),
used in automatic abstraction-based testing.  In TSTL, a test is a
sequence of \emph{actions}; parameter values to method and function
calls of the Software Under Test (SUT) are stored in finite-sized variable \emph{pools}.  A
TSTL test is (essentially) a sequence of assignments to pool values,
and method/function calls using pool values (including method calls on
pool objects).  This formalism is a widely used and efficient way to
represent unit tests \cite{Pacheco,AndrewsTR}.  In TSTL, \emph{visible
value determinism} is based on comparing the values of \emph{all pool
variables} after each test action (pool values other than the one
assigned to must also be compared, since a common source of
nondeterminism is when a call results in a change to a value passed as
a parameter, changes a shared reference held by two values).
TSTL omits comparison only of pools marked as {\tt OPAQUE} in the
TSTL language test harness (and values that do not support equality
checking).  \emph{Final State Determinism} simply performs the same
comparison, but only on the final values of all pool variables (or a
set of designated pools) after a test has
finished executing.
